---
title: "Non-Standard Evaluation"
---

```{r include=FALSE}
library(glue)
library(tidyverse)
library(rlang)
```

Non-standard evaluation is a catch-all term that means they don’t follow the usual R rules
of evaluation. Instead, they capture the expression that you typed and evaluate it in a
custom way.


## Strings and quoting

Here's a simple example in the case of strings. If you want to return a personalised 
message, it is obvious the below function will not give you what you want:

```{r}
greet <- function(name) {
  "Hello, name"
}

greet("Bob")
```

The function returns the string exactly as it is written because it is in quotes. With the 
[`glue`](https://www.rdocumentation.org/packages/glue) package you can use non-standard 
evaluation to return what you want:

```{r}
greet <- function(name) {
  glue("Hello, {name}")
}

greet("Bob")
```

The [`glue()`](https://www.rdocumentation.org/packages/glue/topics/glue) function 
"unquotes" the "name" and evaluates it as a variable.


## Base R evaluation


#### Expressions and quoting

A similar approach can be taken to more general expressions in R. An **expression** is some 
R code has not been evaluated yet. It is captured and saved for later. In base R this can
be achieved using the 
[`quote()`](https://www.rdocumentation.org/packages/base/topics/substitute) function.
Expressions can be one of four classes:

 - "constants" (e.g. 4, TRUE).
 - "names" (e.g. variable names), that have type "symbols".
 - "calls" (e.g. unevaluated function calls), that have type "language".
 - "pairlists" (not really used anymore).

Constants are not very interesting, so we will concentrate on names and calls. As with 
strings, we can capture the expression, rather than the evaluated result, using the 
[`quote()`](https://www.rdocumentation.org/packages/base/topics/substitute) function:

```{r}
quote(x) %>% class()

quote(mean(1:10)) %>% class()
```

To evaluate a quoted expression use the
[`eval()`](https://www.rdocumentation.org/packages/base/topics/eval) function:

```{r}
x <- 10

quote(x) %>% eval()

quote(mean(1:10)) %>% eval()
```


#### Names and Environments

An **environment** is a container for variables, binding a set of names to a set of values.
Every environment also has a parent environment, except the **empty** environment, which is 
the ultimate ancester of all environments. If a name is not found in the current environment
R looks in its parent and so on through the generations. This is known as **lexical
scoping**.

When using `eval()` as above the expression is evaluated in the current environment.
However, the environment to evaluate the expression in can be set as a parameter to
`eval()`.

Note: `data.frames` can be treated as environments containing the column names binded to the
column vectors.


#### Calls

A **call** is a delayed evaluation of a function call. The function and the argument names
are stored, but not evaluated. So you can change the values associated with the arguments
before evaluation. In fact they need not be defined at creation time at all:

```{r}
wait_for_it <- quote(x + y)

x <- 3
y <- 8

eval(wait_for_it)
```


#### Parsing and Deparsing

**Parsing** turns text into expressions and **deparsing** turns expressions into text.
The [`parse()`](https://www.rdocumentation.org/packages/base/topics/parse) function 
converts text, but its default argument is a file connection, so we must use the text
argument:

```{r}
parse(text = "8 + 10") %>% eval()
```

Deparse might be useful for returning information to the user:

```{r}
friendly_eval <- function(expr) {
  str_c("The value of ", deparse(expr), " is ", eval(expr))
}

quote(8 + 10) %>% friendly_eval()
```


#### Functions and closures

When you execute a function it creates a new, temporary, environment. The named arguments 
to the function, plus any variables created within its body are stored in this environment.
Thus it cannot effect the variables outside its scope.

The parent for the function's environment is the environment in which the function was 
created in, not the one in which it was executed in. Thus it has access to all the 
variables in the parent environment. This is known as **closure**, as it _encloses_ the
parent environment, and can be a powerful tool.


#### Substitute and promise

The `friendly_eval()` function above requires its argument to be quoted. It would be nice
if we could write the expression directly as an argument. However, `quote()` makes a 
literal quote of its input, in this case `expr`. What we need is 
[`substitute()`](https://www.rdocumentation.org/packages/base/topics/substitute). This will
lookup all the object names provided to it, and if it finds a value for that name, it will
substitute the name for its value:

```{r}
friendly_eval <- function(expr) {
  expr_sub <- substitute(expr)
  str_c("The value of ", deparse(expr_sub), " is ", eval(expr_sub))
}

friendly_eval(8 + 10)
```

The above function only works because in R function arguments are evaluated lazily --
variables are not evaluated until they are used. R stores arguments as a **promise**, 
which contains the expression of an argument along with the value. `substitute()` 
capture the expression before it is evaluated.


#### Formula and overscoping

A **formula** is a _domain specific language_ (DSL) to simplify expressing the relationship 
between variables. Just like functions, formulas enclose the environment they are created 
in. When a formula is evaluated later in a different environment, it can still access all 
the objects that lived in its original environment.

If an object exists in more than one accessible environment the enviroment the formula
(or function) is evaluated in takes precidence. If the object is not found, R looks 
in the enclosed environment. This is known as **overscoping**, as the formula or 
function has scope beyond its execution environment.


## Tidy Evaluation

To summarise the above, the following points are important to tidy evaluation:

 1. `quote()` delays evaluation of an expression.
 2. `eval()` evaluates an expression in the current (or a specified) environment.

Functions and formulas:

 3. enclose the environment they were created in.
 4. evaluate objects in their own, and enclosed, environments

Tidy evaluation has two new additions: **quasiquotation** and **quosures**.


#### Quasiquotation

**Quasiquotation** enables the user to evaluate parts of the expression right away, while 
quoting the rest. Say we have the expression `z - x + 4` and we know the value of `x` at
the time of quoting, we can **unquote** `x` with the function 
[`UQ()`](http://rlang.tidyverse.org/reference/quasiquotation.html) or the operator
[`!!`](http://rlang.tidyverse.org/reference/quasiquotation.html). However, `quote()` 
will not work in this quasiquotation, we need to use the tidy evaluation equivalent, 
[`expr()`](http://rlang.tidyverse.org/reference/expr.html):

```{r}
x <- 10

expr(z - UQ(x) + 4)
expr(z - !!x + 4) %>% class()
```

The `expr()` function returns expressions, as does `quote()`, without any information on 
the environment it was created in. Some other useful functions for quasiquotation are:

 - [`UQS()`](http://rlang.tidyverse.org/reference/quasiquotation.html) and the 
   [`!!!`](http://rlang.tidyverse.org/reference/quasiquotation.html) operator unquote and 
   splice their arguments.
 - [`enexpr()`](http://rlang.tidyverse.org/reference/expr.html) takes an expression, looks
   up any symbols (names) within it and returns it unevaluated. It is equivalent to
   `substitute()`.
 - [`exprs()`](http://rlang.tidyverse.org/reference/expr.html) captures multiple
   expressions and returns a list.


#### Quosures

As the name suggests, **quosures** are hybrids of quotes and closures. They are 
unevaluated expressions that enclose their creation environment. This is very similar to
formulas and they are in fact implemented as one-sided formulas. Quosures are created with
the [`quo()`](http://rlang.tidyverse.org/reference/quosure.html) function:

```{r}
quo(z - UQ(x) + 4)
quo(z - !!x + 4) %>% class()
```

The quosure equivalent of `substitute()` is
[`enquo()`](http://rlang.tidyverse.org/reference/quosure.html). It takes a symbol referring
to a function argument, quotes the R code that was supplied to this argument, captures the
environment where the function was called (and thus where the R code was typed), and
bundles them in a quosure.

Another useful function is [`quos()`](http://rlang.tidyverse.org/reference/quosure.html), it
returns a list of quosures. You can supply several expressions directly, e.g. 
`quos(foo, bar)`, but more importantly you can also supply dots: `quos(...)`.

Note that quosures don’t make a lower level distinction between calls and names. Every
expression becomes a quosure.

To evaluate quosures we need a special function to implement the environment scoping
properly. `rlang` provides such a function:
[`eval_tidy()`](rlang.tidyverse.org/reference/eval_tidy.html)

```{r}
x <- 10

quo(x) %>% eval_tidy()

quo(mean(1:10)) %>% eval_tidy()
```


#### Parsing and Deparsing

`rlang` also provides functions for parsing and deparsing expressions. To turn a string
into an expression use:
 - [`parse_expr()`](http://rlang.tidyverse.org/reference/parse_expr.html), which works in
   the same way as `parse(text = ...)`.
 - [`parse_quosure()`](http://rlang.tidyverse.org/reference/parse_expr.html) to create a
   quosure.

To turn an expression into a string, `rlang` provides two functions:
 - [`expr_text()`](http://rlang.tidyverse.org/reference/expr_label.html) to turn an
   expression into a string.
 - [`expr_label()`](http://rlang.tidyverse.org/reference/expr_label.html) to produce a
   string that is more suited to use as a label.


```{r}
friendly_eval <- function(expr) {
  str_c("The value of ", deparse(expr), " is ", eval(expr))
}

quote(8 + 10) %>% friendly_eval()
```
